### 2015-11-06 Paper.tex
Final draft

%Original version in April 2002 by Antje Endemann
%
%Adapted for usage in Student Conference course by jubo (050118).
%Optimized for usage with pdflatex. For usage with plain latex,
%figures must be in Postscript (.ps or .eps files) and additional
%packages must be imported.
%Last updated 100105 by jubo
%
%This template works for outline and annotated bibliography
%(deliverable 2) without any changes.
%For usage with full and final papers (deliverables 3 and 4a/b)
%you must make several changes. All necessary changes are
%described in three specific commented sections preceded by
%`%%%%% BEGIN OF ADJUSTMENTS SECTION %%%%%'.
%The changes are in order of appearance:
%   - adjust the `\documentclass' command
%   - uncomment the abstract section
%   - adjust the `\bibliographystyle' command

%%%%% BEGIN OF 1st ADJUSTMENTS SECTION %%%%%
% You must use only one of the following `\documentclass' commands
% For the outline and annotated bibliography (deliverable 2)
% you must use the following command:
%
% \documentclass[runningheads,a4paper,oribibl]{llncs}
%
% For the full and final paper (deliverables 3 and 4a/b)
% you must use the command below (and put the one above in comments).
% In other words, the 'oribib' option must be used for Deliverable 2
% but not for 3 and 4a/b.
% 
%
\documentclass[runningheads,a4paper]{llncs}
%
%%%%% END OF 1st ADJUSTMENTS SECTION %%%%%

\usepackage[T1]{fontenc}  %% needed for special characters (umlaut)
\usepackage[utf8]{inputenc}
\usepackage{graphicx}     %% for graphical things such as including pictures
\usepackage{url}          %% for proper formatting of URLs

% Albins Packages
\usepackage{array}

\begin{document}

\pagestyle{headings}

\mainmatter

% \title{Guidelines for Designing Touch Interfaces for Controlling Robotic Nozzles in Critical Emergency Situations.}

\title{Can a High Color Contrast Touch Interface Increase User Reaction Time when Using a Smart Phone Web Based Application.}

% response
% reaction time

% The abbreviated title will be shown in the headers of even pages.
% You should use the full title unless it is too long.

% \titlerunning{Touch Interfaces for Robotic Nozzles in Emergency Situations}

\titlerunning{Can High Color Contrast Increase User Reaction Time}

\author{Albin Hübsch}

\institute{
	Department of Computing Science \\
	Umeå University, Sweden \\
	\email{id11ahh@cs.umu.se}
}

\maketitle

%%%%% BEGIN OF 2nd ADJUSTMENTS SECTION %%%%%
% Do NOT include an abstract in the outline and annotated bibliography
% (deliverable 2). For the full and final paper (deliverables 3 and 4a/b),
% you must uncomment the following two commands and place your text for
% the abstract in between them.
%
\begin{abstract}
Here goes paper abstract
\end{abstract}
%
%%%%% END OF 2nd ADJUSTMENTS SECTION %%%%%

\section{Introduction}
The main goal with this paper is to evaluate if color contrast has a significant impact on usability in web based smart-phone applications. Consumer touch screen devices such as smart-phones has rapidly increased in amount and availability recent years. The touch screen technology have made great advances ~\cite{jennings2013touch} and it is more frequently used as a way of receiving user input. When this technology moves to a broader audience, higher demands on usability needs to be set~\cite{gong2004guidelines}. To further explore how usability can be improved using touch screen technology we are in this paper investigating if reaction time (referred to as RT\footnote{Reaction Time: the time it takes for a user to react on instructions} in this paper) and user input errors (referred to as IE\footnote{Input Errors: the number of user input errors when interacting with a system} in this paper) can reach better results by increasing color contrast within the interface. Earlier work done in this field are exploring how we perceive different color contrast and how it can affect our reading performance~\cite{wu2003improving}. It has also been proved that Chromaticity, Contrast, and Cone Opponency in color space can affect RTs~\cite{mckeefry2003simple}. Many best practices for Mobile development also shows that "requirements for sufficient color contrast"~\cite{marcus2013design} must be meet to best exaggerate the content. But still the question remains, if a high color contrast interface can increase the the RT and decrease user IE.

High RTs and low user IE are especially desirable when designing user interfaces for situations with high demands on quick user input and low error tolerance such as emergency situations. The results of this paper can be used by any designer or developer.

\section{Method}
In order to be able to test if color contrast has an substantial impact on the RT and user IE of a touch user interface we have designed a simple web application for an Android smart-phone. The application exists in two versions, one with low color contrast (referred to as LC in this paper) and one with high color contrast (referred to as HC in this paper) as can be seen in Fig~\ref{fig:application}.

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{application}
	\caption{Our two versions of the application designed with different color contrast.
	\label{fig:application}}
\end{figure}

We designed the application by following design guidelines given here~\cite{hoober2011designing}~\cite{johnson2013designing}~\cite{gong2004guidelines}~\cite{norman2013design}.

\subsection{Designing the A/B test}
We conducted an A/B test to measure the differences in RT and IE between the two versions. An A/B test is a commonly used method by developers and designers to test differences in performance between applications when the differentiated factors are known~\cite{johnson2013designing}. As in this case our known factor is the contrast difference.

The application consists of 4 buttons\footnote{On screen button}, each button representing a function. To fraction each button from the others we put a unique shape in each button. The shapes we used are a, square, circle, triangle and a rhomb~\ref{fig:application}.

To be able to give the test persons consistent instructions through the complete test we designed a program that presented the instructions on a secondary screen. 
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{instructions}
	\caption{Our program designed to present user instructions.
	\label{fig:instructions}}
\end{figure}
The instructions was a series of 20 instructions with shapes equally to the shapes in the application buttons. The test persons where told to press the representative shape in the application as fast as they could to measure the reaction time. If they pressed the wrong button they where told to continue the test by pressing the right button, the application registered this as an IE. Each test were performed indoors with varying surroundings such as different lightning conditions and noise levels. We tested each version of the application on 5 persons all within the same age group (20-30) and with a variety of backgrounds. We had a gender split of 50\% women and 50\% men. In total we got 10 test results that we later grouped into LC and HC. We analyzed them against each other to measure if any differences in RTs or IEs between the two versions could be statistically confirmed.

\subsection{Evaluation of the A/B test}
Our test data received from the tests consisted of UNIX timestamps\footnote{\url{http://www.unixtimestamp.com/}} in milliseconds. Each time the user interacted with the application it registered which button that was pressed and saved it together with a UNIX timestamp. Equally, every time the instruction program showed a new symbol it logged that together with a UNIX timestamp. As a result we received two data sets of symbols with corresponding timestamps for each test person, resulting in a total of approximately 400 records. Although the data were hard to interpret due to its compact look, a python script were written to clean it up. We calculated each difference (RT) in time between when the instruction were given and when user interaction registered. Out of this we could calculate the mean RT for both groups LC and HC. The calculation of RT did not take in consideration if user input were a no match. We still considered it as an reaction.

IEs were counted every time the user pressed a non matching object compared to the one given in the instructions. All IEs in each group, LC and HC, were counted to see if one of the groups produced more IEs.

\section{Result}
From our results we can not draw any statistically significant differences in RT between the two groups HC and LC. This is most probably due to corrupt test data, a result from limitations in web application technology, see section \ref{subsec:discussion}. With IEs we found that the LC group produced significantly more IEs than the HC group.

\subsection{Reaction Times (RT)}
Table~\ref{tab:groupRT} shows our average RT and standard deviation for each group. According to the results people in HC had insignificant slower RTs compared to the ones in LC.

\begin{table}[]
	\centering
	\label{tab:groupRT}
	\setlength{\tabcolsep}{1em}
	\setlength\extrarowheight{1em}
	\begin{tabular}{l|l|l}
		\textbf{} & \textbf{Average RT} & \textbf{Standard Deviation} \\ \hline
		\textbf{HC} & 782.69 ms & 968.67 ms \\ \hline
		\textbf{LC} & 549.52 ms & 1393.03 ms
	\end{tabular}
	\caption{Average RT and standard deviation for each group, LC and HC.}
\end{table}

To validate that the groups RTs are insignificantly different we assumed the two data sets where normal distributed and performed a two tailed T-Test. The T-Test produced a p-value of 0.17 which is bigger than a desired error margin ($\alpha$) of 5\%. The T-Test is therefore proving that no differences in RT can be shown with our data. But regardless what the t-test says our standard deviations are hinting that something might be wrong with our data, please see section \ref{subsec:discussion}.

\subsection{Input Errors (IE)}
None of the groups where completely free from user IEs. As can be seen in table~\ref{tab:userIE} users that were given the LC interface produced 475\% more IEs than the HC group.

\begin{table}[]
	\centering
	\label{tab:userIE}
	\setlength{\tabcolsep}{1em}
	\setlength\extrarowheight{1em}
	\begin{tabular}{l|l|l|l}
		\textbf{} & \textbf{Errors} & \textbf{Standard Deviation} & \textbf{Errors/User} \\ \hline
		\textbf{HC} & 4 & 1.10 & 0.8 \\ \hline
		\textbf{LC} & 19 & 3.49 & 3.8
	\end{tabular}
	\caption{Number of user IEs registered in each group and a calculation of IEs per user in each group}
\end{table}

Table~\ref{tab:userIE} shows us that there is a significant difference in number of IEs between the groups. The results also show us that the users in the HC group generally performed more stable and consistent compared to the LC group.

\section{Discussion}\label{sec:discussion}
Our test results has shown that we can not increase the user reaction time by using a high color contrast interface when using a smart phone web based application. However we could see that the high color contrast interface was more reliable in performance when it comes to the amount of user input errors. This is fact according to our results. But what we have seen in our collected test data can strongly negotiate the legibility of the results. The calculated standard deviations gave us hints on that something were wrong so we dug deeper into each test persons results and found that the data must be corrupt. We found that we randomly receive negative RTs which in theory should be completely impossible. We were prepared that the timestamps could be off sync with each others but that would be easy to compensate. Here we got inconsistent hick ups in our timestamps. We suspect this is a result from either hardware limitations or limitations in the JavaScript\footnote{A prototype based dynamic scripting language} engine for web applications on android.

As for this our conclusion is that HC interfaces are the preferred method to design interfaces as they deliver a more stable user interaction pattern. Apart from this, smart phone web based applications should only be used if, sometimes slow technology, user input errors and long reaction times from the user can be accepted.

\subsection{Drawbacks and Limitations}\label{subsec:drawbacks}
The main drawback of this study is the corrupt data set. We have to have this in mind when we draw any conclusions out of the results. This problem could be solved by simply rerun the whole test using a native smart phone application instead of a web based solution.

Another drawback worth to mention is that, preferable we should also have had let the test persons do the test several number of times in order to eliminate possible surrounding factors that could have affected the results. We should also had let some of the test persons perform tests with both the LC and HC versions in order to see how fast they memorize patterns.

Further are some more drawbacks and limitations that we would like to mention.
\begin{description}
	\item[Our group of test subjects is to small] To be able to get a more statistical valid result the test group has to be bigger. In this paper we also only assumed it was normal distributed, not tested.
	\item[Many affecting parameters] The prototype is simple in its appearance but there are still many parameters that can affect the users interaction and their RTs. Icons, icon size, optimal number of soft-buttons, size of hand-held device etc. All this, and more, should have been researched and taken into account if done again. 
	\item[Limited test environment] Our tests were made using an Android based smart-phone\footnote{OnePlus two \url{https://oneplus.net/2}} system and the application were made using HTML5\footnote{HyperText Markup Language, markup standard for WWW}, CSS3\footnote{Cascading Style Sheet, styling standard for WWW} and JavaScript. We can not surely imply, without testing, that the same results would appear on an iPhone or any other smart-phone or even with a native application.
\end{description}

\subsection{Future Work}
Due to the failed data set that we received in this paper there are things that could be improved or continued to be worked on. 
We do have some concrete suggestions that could be case for future work.

\begin{itemize}
	\item Do tests using a native application.
	\item Extend the test group with more test subjects. It is possible to think that the patterns that slightly appeared in our results will appear even more significant with a larger group of test subjects.
	\item Our collected test data can be downloaded and used freely to investigate other aspects not mentioned in this paper. One proposal is to look at the different shapes and try to detect possible error patterns between them. Is the rhomb more frequently a case for IE compared to the other shapes?
	\item Does our results apply on other systems and techniques? Our tests are limited to the Android system and a web based application solution. It remains to answer if our results applies on all other systems. This is also a way of isolating some of the affecting parameters mentioned in~\ref{subsec:drawbacks}.
\end{itemize}

\section{Acknowledgments}
The author would like to thank the peer reviewers who provided great valuable comments on both the content and structure of this paper.

%%%%% BEGIN OF 3rd ADJUSTMENTS SECTION %%%%%
% For the outline and annotated bibliography (deliverable 2)
% you must use the following two commands:
%
% \nocite{*}  % Includes ALL entries from the .bib-file, even if they are
            % not '\cite{}'-ed in the text above
% \bibliographystyle{plain-annote}
%
% For the full and final paper (deliverables 3 and 4a/b)
% use ONLY the following commands (i.e. put the commands above into
% comments and uncomment the command below):
%
\bibliographystyle{splncs}
%
%%%%% END OF ADJUSTMENTS SECTION %%%%%

\bibliography{Bibliography}

\end{document}